Optimal design of experiment

### 2025-05-21 Article_Design_of_Experiment-Chamomile.tex
21.05

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study investigates the process of chamomile oil extraction from flowers. A parameter-distributed model consisting of a set of partial differential equations is used to describe the governing mass transfer phenomena in a{ \color{blue}cylindrical packed bed with solid chamomile particles} under supercritical conditions using carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. The empirical correlations used in the model are based on the set of laboratory experiments performed under multiple constant operating conditions: ${\color{blue}31} - 40~^\circ$C, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. A model-based design of experiments with D-optimality criterion is applied to improve the precision of the correlation parameters by designing a new dynamic experiment. The mass flow rate and inlet temperature are used as decision variables to maximize the Fisher Information embedded in the yield curve with respect to the empirical correlations.

@@

\nomenclature[A]{\(P\)}{Pressure \nomunit{bar}}
\nomenclature[A]{\(T\)}{Temperature \nomunit{K}}
\nomenclature[A]{\(T_{in}\)}{Inlet temperature \nomunit{K}}
\nomenclature[A]{\(T_{out}\)}{Outlet temperature \nomunit{K}}
\nomenclature[A]{\(F\)}{Mass flow rate \nomunit{kg/s}}
\nomenclature[A]{\(u\)}{Superficial velocity \nomunit{m/s}}
\nomenclature[A]{\(v\)}{Linear velocity \nomunit{m/s}}
\nomenclature[A]{\(z\)}{Spatial direction \nomunit{m}}
\nomenclature[A]{\(e\)}{Internal energy \nomunit{J/kg}}
%\nomenclature[A]{\(e^y\)}{Output error}
\nomenclature[A]{\(h\)}{Enthalpy \nomunit{kJ/kg}}
\nomenclature[A]{\(t\)}{Time \nomunit{s}}
\nomenclature[A]{\(t_f\)}{Total extraction time \nomunit{s}}
\nomenclature[A]{\(t_0\)}{Initial extraction time \nomunit{s}}
\nomenclature[A]{\(A\)}{Total cross-section of the bed \nomunit{m$^2$}}
\nomenclature[A]{\(A_f\)}{Cross-section of the bed occupied by the fluid \nomunit{m$^2$}}
\nomenclature[A]{\(c_f\)}{Concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{f0}\)}{Initial concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_f^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s\)}{Concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{s0}\)}{Initial concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_p\)}{Concentration of solute in the core of a pore \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{pf}\)}{Concentration of solute at the pore opening \nomunit{kg/m$^3$}}
\nomenclature[A]{\(D_e^M\)}{Axial diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r_e\)}{Mass transfer kinetic term \nomunit{kg/m$^3$/s}}
\nomenclature[A]{\(L\)}{Length of the fixed bed \nomunit{m}}
\nomenclature[A]{\(j\)}{Objective function \nomunit{-}}
\nomenclature[A]{\(j_{ml}\)}{Maximum Likelihood \nomunit{-}}
\nomenclature[A]{\(S\)}{Sensitivity equations \nomunit{-}}
\nomenclature[A]{\(Q\)}{Weighting matrix \nomunit{-}}
\nomenclature[A]{\(R\)}{Control cost matrix \nomunit{-}}
%\nomenclature[A]{\(\textbf{G}\)}{Augumented system}
%\nomenclature[A]{\(\dot{S}\)}{Time derivative of sensitivity equations}
%\nomenclature[A]{\(\bar{S}\)}{Sensitivity matrix}
%\nomenclature[A]{\(\bar{J}_x\)}{Jacobian matrix with respect to the state space}
%\nomenclature[A]{\(\bar{J}_\Theta\)}{Jacobian matrix with respect to the paramters}
%\nomenclature[A]{\(\mathcal{L}\)}{Log-likelihood function}
\nomenclature[A]{\(l\)}{Characteristic dimension of particles \nomunit{m}}
\nomenclature[A]{\(D_i\)}{Internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(D_i^R\)}{Reference value of internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r\)}{Particle radius \nomunit{m}}
\nomenclature[A]{\(k_p\)}{Volumetric partition coefficient \nomunit{-}}
\nomenclature[A]{\(k_m\)}{Mass partition coefficient \nomunit{-}}
\nomenclature[A]{\(y\)}{Extraction yield \nomunit{g}}
%\nomenclature[A]{\(Y\)}{Yield measurement \nomunit{g}}
\nomenclature[A]{\(x\)}{State vector}
\nomenclature[A]{\(G\)}{Vector of discretized differential equations}
\nomenclature[A]{\(p\)}{Probaility disribution model}
\nomenclature[A]{\(R_e\)}{Reynolds number}
\nomenclature[A]{\(\mathcal{F}\)}{Fisher information}

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
This study investigates the process of chamomile oil extraction from flowers. A parameter-distributed model consisting of a set of partial differential equations is used to describe the governing mass transfer phenomena in a cylindrical packed bed with solid chamomile particles under supercritical conditions using carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. The empirical correlations used in the model are based on the set of laboratory experiments performed under multiple constant operating conditions: $30 - 40~^\circ$C, $100 - 200$ bar, and $3.33-6.67 \cdot 10^{-5}$ kg/s. A model-based design of experiments with D-optimality criterion is applied to improve the precision of the correlation parameters by designing a new dynamic experiment. The mass flow rate and inlet temperature are used as decision variables to maximize the Fisher Information embedded in the yield curve with respect to the empirical correlations.

@@

\nomenclature[A]{\(P\)}{Pressure \nomunit{bar}}
\nomenclature[A]{\(T\)}{Temperature \nomunit{K}}
\nomenclature[A]{\(T_{in}\)}{Inlet temperature \nomunit{K}}
\nomenclature[A]{\(T_{out}\)}{Outlet temperature \nomunit{K}}
\nomenclature[A]{\(F\)}{Mass flow rate \nomunit{kg/s}}
\nomenclature[A]{\(u\)}{Superficial velocity \nomunit{m/s}}
\nomenclature[A]{\(v\)}{Linear velocity \nomunit{m/s}}
\nomenclature[A]{\(z\)}{Spatial direction \nomunit{m}}
\nomenclature[A]{\(e\)}{Internal energy \nomunit{J/kg}}
%\nomenclature[A]{\(e^y\)}{Output error}
\nomenclature[A]{\(h\)}{Enthalpy \nomunit{kJ/kg}}
\nomenclature[A]{\(t\)}{Time \nomunit{s}}
\nomenclature[A]{\(t_f\)}{Total extraction time \nomunit{s}}
\nomenclature[A]{\(t_0\)}{Initial extraction time \nomunit{s}}
\nomenclature[A]{\(A\)}{Total cross-section of the bed \nomunit{m$^2$}}
\nomenclature[A]{\(A_f\)}{Cross-section of the bed occupied by the fluid \nomunit{m$^2$}}
\nomenclature[A]{\(c_f\)}{Concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{f0}\)}{Initial concentration of solute in the fluid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_f^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s\)}{Concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{s0}\)}{Initial concentration of solute in the solid phase \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_s^*\)}{Concentration of solute at the solid-fluid interface \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_p\)}{Concentration of solute in the core of a pore \nomunit{kg/m$^3$}}
\nomenclature[A]{\(c_{pf}\)}{Concentration of solute at the pore opening \nomunit{kg/m$^3$}}
\nomenclature[A]{\(D_e^M\)}{Axial diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r_e\)}{Mass transfer kinetic term \nomunit{kg/m$^3$/s}}
\nomenclature[A]{\(L\)}{Length of the fixed bed \nomunit{m}}
\nomenclature[A]{\(j\)}{Objective function \nomunit{-}}
\nomenclature[A]{\(j_{ml}\)}{Maximum Likelihood \nomunit{-}}
\nomenclature[A]{\(S\)}{Sensitivity equations \nomunit{-}}
\nomenclature[A]{\(Q\)}{Weighting matrix \nomunit{-}}
\nomenclature[A]{\(\mathcal{R}\)}{Control cost matrix \nomunit{-}}
%\nomenclature[A]{\(\textbf{G}\)}{Augumented system}
%\nomenclature[A]{\(\dot{S}\)}{Time derivative of sensitivity equations}
%\nomenclature[A]{\(\bar{S}\)}{Sensitivity matrix}
%\nomenclature[A]{\(\bar{J}_x\)}{Jacobian matrix with respect to the state space}
%\nomenclature[A]{\(\bar{J}_\Theta\)}{Jacobian matrix with respect to the paramters}
%\nomenclature[A]{\(\mathcal{L}\)}{Log-likelihood function}
\nomenclature[A]{\(l\)}{Characteristic dimension of particles \nomunit{m}}
\nomenclature[A]{\(D_i\)}{Internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(D_i^R\)}{Reference value of internal diffusion coefficient \nomunit{m$^2$/s}}
\nomenclature[A]{\(r\)}{Particle radius \nomunit{m}}
\nomenclature[A]{\(k_p\)}{Volumetric partition coefficient \nomunit{-}}
\nomenclature[A]{\(k_m\)}{Mass partition coefficient \nomunit{-}}
\nomenclature[A]{\(y\)}{Extraction yield \nomunit{g}}
%\nomenclature[A]{\(Y\)}{Yield measurement \nomunit{g}}
\nomenclature[A]{\(x\)}{State vector}
\nomenclature[A]{\(G\)}{Vector of discretized differential equations}
\nomenclature[A]{\(p\)}{Probaility disribution model}
\nomenclature[A]{\(R_e\)}{Reynolds number}
\nomenclature[A]{\(\mathcal{F}\)}{Fisher information}

### 2025-05-21 Sections/DOE.tex
21.05

	{\footnotesize
		\begin{flalign*}
			\ln p\left(Y \mid \theta, \Xi \right) &\propto -\frac{1}{2} \left( \underbrace{r^\top \Sigma_Y^{-1} r}_{C} - 2 \underbrace{ J^\top \Sigma_Y^{-1} \Delta_y }_{B} \Delta_\theta + \Delta_\theta^\top \underbrace{\left( J^\top \Sigma_Y^{-1} J + \Sigma_\theta^{-1} \right)}_{A} \Delta_\theta \right) && \\
			&=-\frac{1}{2} \left( \Delta_\theta^\top A \Delta_\theta - 2 B^\top \Delta_\theta + C \right) &&
		\end{flalign*}
	}

	By completing the square, we introduce $\theta^* = \hat{\theta} + A^{-1} B$, and the cost function becomes:

	{\footnotesize 
		\begin{equation} 
			\ln p\left(\theta \mid Y, \Xi \right) \propto -\frac{1}{2} \left( \left( \theta - \theta^* \right)^\top A \left( \theta - \theta^* \right) + C - B^\top A^{-1} B \right) 
		\end{equation} }

	It can be observed that the quadratic term represents the spread of the posterior multivariate distribution around the mean $\theta^*$. The constant term $C - B^\top A^{-1} B$ is independent of $\theta$ and does not affect the shape of the posterior distribution but is still important for normalization. This constant term is absorbed into the normalizing constant of the posterior distribution. It does not affect the estimation of $\theta$ or the shape of the posterior distribution and thus can be neglected when focusing on parameter estimation or experimental design. Therefore, the posterior distribution of $\theta$ given $Y$ and $\Xi$ is:

	{\footnotesize \begin{equation} p\left(\theta \mid Y, \Xi \right) \propto \exp \left( -\frac{1}{2} \left( \theta - \theta^* \right)^\top A \left( \theta - \theta^* \right) \right) \end{equation} }

	This is the expression of a multivariate normal distribution with mean $\theta^*$ and covariance matrix $A^{-1}$:

	{\footnotesize \begin{equation} p\left(\theta \mid Y, \Xi \right) = \mathcal{N}\left( \theta^*, A^{-1} \right) \end{equation} }

	As the goal of optimal experimental design is to improve the precision of parameters, the main focus is on the matrix $A$, which consists of two terms. The first corresponds to the information gained from new observations based on the model and measurement error ($J^\top \Sigma_Y^{-1} J$), while the second is the prior information ($\Sigma_\theta^{-1}$).

@@

	The cost function should describe the amount of information obtained from an experiment. For that purpose, it can be assumed that a function $\kappa$ can be related to the covariance matrix $A^{-1}$ obtained at arbitrary operating conditions:

	{\footnotesize \begin{equation} j(\Xi) = \kappa\left[ A^{-1}(\theta, \Xi) \right] \end{equation} }

	Recall that in the Bayesian estimation framework, the posterior covariance matrix of the parameter estimates is given by $\Sigma = A^{-1} = \left(J^\top \Sigma_Y^{-1} J + \Sigma_\theta^{-1}\right)^{-1} $. $J$ is the Jacobian matrix of the model outputs with respect to the parameters, $\Sigma_Y$ is the covariance matrix of the measurement errors, and $\Sigma_\theta$ is the prior covariance matrix of the parameters.

@@

	{\footnotesize 
		\begin{align} \label{EQ:DOE_general}
			\kappa_k\left( A^{-1}(\theta) \right) &= \left[ \frac{1}{n_\theta} \operatorname{trace}\left( Q A^{-1}(\theta, \Xi) Q^\top \right)^k \right]^{1/k} & \text{if} \det A^{-1} \neq 0, \nonumber \\ 
			\kappa_k\left( A^{-1}(\theta) \right) &= \infty & \text{if} \det A^{-1} = 0
		\end{align} }

@@

	{\footnotesize \begin{equation} j_L(\Xi) = \operatorname{trace} \left( Q A^{-1}(\theta, \Xi) Q^\top \right) \end{equation} }

@@

	Selecting $Q$ as a row vector results in C-optimality, which focuses on linear combinations of parameters. Setting $Q = \mathbf{I}_{n\theta}$ and $k = \infty$ leads to E-optimality, where the design maximizes the smallest eigenvalue of the covariance matrix $A^{-1}(\theta, \Xi)$, thereby minimizing the length of the largest axis of the asymptotic confidence ellipsoids.

	The most commonly used criterion is the D-optimality, which involves $k = 0$ and $Q = \mathbf{I}_{n\theta}$, requiring the minimization of $\det A^{-1}(\theta, \Xi)$, or equivalently, maximization of:

	{\footnotesize \begin{equation} j_D(\Xi) = \det A(\theta, \Xi) \end{equation} }

	Since $\det A(\theta, \Xi) = 1 / \det A^{-1}(\theta, \Xi)$, maximizing $\det A$ is equivalent to minimizing $\det A^{-1}$.

@@

	Details on the process model can be found in \citet{Sliczniuk2024}. The model's empirical correlations are derived based on laboratory experiments conducted under various but constant operating conditions: $31 - 40~^\circ\text{C}$, $100 - 200$ bar, and $3.33 \cdot 10^{-5} - 6.67 \cdot 10^{-5}$ kg/s. This study employs the process model obtained previously to design a set of supplementary experiments with dynamically changing conditions to improve the precision of the correlation for $D_i$. The analyzed correlations consist of three independent variables: mass flow rate, pressure and inlet temperature.

@@

	In a real system, the mass flow rate and inlet temperature cannot be changed instantaneously to any arbitrary value, as they depend on the dynamics of the pump and heat exchanger, respectively. To prevent bang-bang type of control, a penalty term is introduced. Inspired by control problems, such as the Linear Quadratic Regulator, a quadratic penalty term is added to the cost function. This penalty increases the cost function value when there are rapid changes in the decision variables. The matrix $\bar{R}$ represents the control cost matrix, with its entries chosen so that the control costs are one order of magnitude lower than $-\ln j_D$. The problem formulation is given by Equation \ref{EQ:Formulation_2}.

	{\footnotesize 
		\begin{equation}
			\begin{aligned}
				\Xi^* &= \arg \min_{T^{\text{in}}(t), F(t)} 
				\int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + 
				\frac{du(t)^\top}{dt} \, \bar{R} \, \frac{du(t)}{dt} \right] dt, \\
				\text{subject to:} \quad 
				& \dot{x} = G(x, t, \theta; \Xi), \\
				& t_0 = 0 \, \text{min}, \quad t_f = 600 \, \text{min}, \\
				& T^0 = T^{\text{in}}(t=0), \\
				& P(t) \in \{100, 125, 150, 175, 200\} \, \text{bar}, \\
				& 31^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\
				& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5}, \\
				& \bar{R} = 
				\begin{bmatrix} 
					0.1 & 0 \\ 
					0 & 0.01 
				\end{bmatrix}, \\
				& \frac{du(t)}{dt} = 
				\begin{bmatrix} 
					\dfrac{dF(t)}{dt} \\ 
					\dfrac{dT^{\text{in}}(t)}{dt} 
				\end{bmatrix}.
			\end{aligned}
			\label{EQ:Formulation_2}
			\end{equation} }

>>>>

	{\footnotesize
		\begin{flalign*}
			\ln p\left(Y \mid \theta, \Xi \right) &\propto -\frac{1}{2} \left( \underbrace{r^\top \Sigma_Y^{-1} r}_{\mathcal{C}} - 2 \underbrace{ J^\top \Sigma_Y^{-1} \Delta_y }_{\mathcal{B}} \Delta_\theta + \Delta_\theta^\top \underbrace{\left( J^\top \Sigma_Y^{-1} J + \Sigma_\theta^{-1} \right)}_{\mathcal{A}} \Delta_\theta \right) && \\
			&=-\frac{1}{2} \left( \Delta_\theta^\top \mathcal{A} \Delta_\theta - 2 \mathcal{B}^\top \Delta_\theta + \mathcal{C} \right) &&
		\end{flalign*}
	}

	By completing the square, we introduce $\theta^* = \hat{\theta} + \mathcal{A}^{-1} \mathcal{B}$, and the cost function becomes:

	{\footnotesize 
		\begin{equation} 
			\ln p\left(\theta \mid Y, \Xi \right) \propto -\frac{1}{2} \left( \left( \theta - \theta^* \right)^\top \mathcal{A} \left( \theta - \theta^* \right) + \mathcal{C} - \mathcal{B}^\top \mathcal{A}^{-1} \mathcal{B} \right) 
		\end{equation} }

	It can be observed that the quadratic term represents the spread of the posterior multivariate distribution around the mean $\theta^*$. The constant term $\mathcal{C} - \mathcal{B}^\top \mathcal{A}^{-1} \mathcal{B}$ is independent of $\theta$ and does not affect the shape of the posterior distribution but is still important for normalization. This constant term is absorbed into the normalizing constant of the posterior distribution. It does not affect the estimation of $\theta$ or the shape of the posterior distribution and thus can be neglected when focusing on parameter estimation or experimental design. Therefore, the posterior distribution of $\theta$ given $Y$ and $\Xi$ is:

	{\footnotesize \begin{equation} p\left(\theta \mid Y, \Xi \right) \propto \exp \left( -\frac{1}{2} \left( \theta - \theta^* \right)^\top \mathcal{A} \left( \theta - \theta^* \right) \right) \end{equation} }

	This is the expression of a multivariate normal distribution with mean $\theta^*$ and covariance matrix $\mathcal{A}^{-1}$:

	{\footnotesize \begin{equation} p\left(\theta \mid Y, \Xi \right) = \mathcal{N}\left( \theta^*, \mathcal{A}^{-1} \right) \end{equation} }

	As the goal of optimal experimental design is to improve the precision of parameters, the main focus is on the matrix $\mathcal{A}$, which consists of two terms. The first corresponds to the information gained from new observations based on the model and measurement error ($J^\top \Sigma_Y^{-1} J$), while the second is the prior information ($\Sigma_\theta^{-1}$).

@@

	The cost function should describe the amount of information obtained from an experiment. For that purpose, it can be assumed that a function $\kappa$ can be related to the covariance matrix $\mathcal{A}^{-1}$ obtained at arbitrary operating conditions:

	{\footnotesize \begin{equation} j(\Xi) = \kappa\left[ \mathcal{A}^{-1}(\theta, \Xi) \right] \end{equation} }

	Recall that in the Bayesian estimation framework, the posterior covariance matrix of the parameter estimates is given by $\Sigma = \mathcal{A}^{-1} = \left(J^\top \Sigma_Y^{-1} J + \Sigma_\theta^{-1}\right)^{-1} $. $J$ is the Jacobian matrix of the model outputs with respect to the parameters, $\Sigma_Y$ is the covariance matrix of the measurement errors, and $\Sigma_\theta$ is the prior covariance matrix of the parameters.

@@

	{\footnotesize 
		\begin{align} \label{EQ:DOE_general}
			\kappa_k\left( \mathcal{A}^{-1}(\theta) \right) &= \left[ \frac{1}{n_\theta} \operatorname{trace}\left( Q \mathcal{A}^{-1}(\theta, \Xi) Q^\top \right)^k \right]^{1/k} & \text{if} \det \mathcal{A}^{-1} \neq 0, \nonumber \\ 
			\kappa_k\left( \mathcal{A}^{-1}(\theta) \right) &= \infty & \text{if} \det \mathcal{A}^{-1} = 0
		\end{align} }

@@

	{\footnotesize \begin{equation} j_L(\Xi) = \operatorname{trace} \left( Q \mathcal{A}^{-1}(\theta, \Xi) Q^\top \right) \end{equation} }

@@

	Selecting $Q$ as a row vector results in C-optimality, which focuses on linear combinations of parameters. Setting $Q = \mathbf{I}_{n\theta}$ and $k = \infty$ leads to E-optimality, where the design maximizes the smallest eigenvalue of the covariance matrix $\mathcal{A}^{-1}(\theta, \Xi)$, thereby minimizing the length of the largest axis of the asymptotic confidence ellipsoids.

	The most commonly used criterion is the D-optimality, which involves $k = 0$ and $Q = \mathbf{I}_{n\theta}$, requiring the minimization of $\det \mathcal{A}^{-1}(\theta, \Xi)$, or equivalently, maximization of:

	{\footnotesize \begin{equation} j_D(\Xi) = \det \mathcal{A}(\theta, \Xi) \end{equation} }

	Since $\det A(\theta, \Xi) = 1 / \det \mathcal{A}^{-1}(\theta, \Xi)$, maximizing $\det \mathcal{A}$ is equivalent to minimizing $\det \mathcal{A}^{-1}$.

@@

	Details on the process model can be found in \citet{Sliczniuk2024}. The model's empirical correlations are derived based on laboratory experiments conducted under various but constant operating conditions: $30 - 40~^\circ\text{C}$, $100 - 200$ bar, and $3.33 \cdot 10^{-5} - 6.67 \cdot 10^{-5}$ kg/s. This study employs the process model obtained previously to design a set of supplementary experiments with dynamically changing conditions to improve the precision of the correlation for $D_i$. The analyzed correlations consist of three independent variables: mass flow rate, pressure and inlet temperature.

@@

	In a real system, the mass flow rate and inlet temperature cannot be changed instantaneously to any arbitrary value, as they depend on the dynamics of the pump and heat exchanger, respectively. To prevent bang-bang type of control, a penalty term is introduced. Inspired by control problems, such as the Linear Quadratic Regulator, a quadratic penalty term is added to the cost function. This penalty increases the cost function value when there are rapid changes in the decision variables. The matrix $\mathcal{R}$ represents the control cost matrix, with its entries chosen so that the control costs are one order of magnitude lower than $-\ln j_D$. The problem formulation is given by Equation \ref{EQ:Formulation_2}.

	{\footnotesize 
		\begin{equation}
			\begin{aligned}
				\Xi^* &= \arg \min_{T^{\text{in}}(t), F(t)} 
				\int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + 
				\frac{du(t)^\top}{dt} \, \bar{R} \, \frac{du(t)}{dt} \right] dt, \\
				\text{subject to:} \quad 
				& \dot{x} = G(x, t, \theta; \Xi), \\
				& t_0 = 0 \, \text{min}, \quad t_f = 600 \, \text{min}, \\
				& T^0 = T^{\text{in}}(t=0), \\
				& P(t) \in \{100, 125, 150, 175, 200\} \, \text{bar}, \\
				& 31^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\
				& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5}, \\
				& \mathcal{R} = 
				\begin{bmatrix} 
					0.1 & 0 \\ 
					0 & 0.01 
				\end{bmatrix}, \\
				& \frac{du(t)}{dt} = 
				\begin{bmatrix} 
					\dfrac{dF(t)}{dt} \\ 
					\dfrac{dT^{\text{in}}(t)}{dt} 
				\end{bmatrix}.
			\end{aligned}
			\label{EQ:Formulation_2}
			\end{equation} }

### 2025-05-21 Sections/Model.tex
21.05

			{\color{blue}As explained in Section \ref{CH:Governing_equations_chapter}, the pressure in the low-velocity region remains nearly constant due to the small pressure wave propagation occurring at the speed of sound. Under such conditions, the term $\partial P/\partial t$ can be approximated by a forward difference equation, describing the pressure change over a small time step $\Delta t$. The pressure $P$ in the system is treated as a state variable, while the incoming pressure at the new time-step $P_{in}$ is treated as a control:}

@@

			{\color{blue}Such a simplified equation allows for instantaneous pressure. In a real system, the pressure dynamics would depend on the behaviour of the pump and the back-pressure regulator, which introduce additional inertia and resistance to pressure changes, leading to pressure gradual build-up.}

>>>>

			As explained in Section \ref{CH:Governing_equations_chapter}, the pressure in the low-velocity region remains nearly constant due to the small pressure wave propagation occurring at the speed of sound. Under such conditions, the term $\partial P/\partial t$ can be approximated by a forward difference equation, describing the pressure change over a small time step $\Delta t$. The pressure $P$ in the system is treated as a state variable, while the incoming pressure at the new time-step $P_{in}$ is treated as a control:

@@

			Such a simplified equation allows for instantaneous pressure. In a real system, the pressure dynamics would depend on the behaviour of the pump and the back-pressure regulator, which introduce additional inertia and resistance to pressure changes, leading to pressure gradual build-up.
